---
title: "Course 5 - Project Assignment 1 - Final Version"
author: "Pedro Magalhaes Batista"
date: "12/11/2018"
output: html_document
---
# SUMMARY
1. Parsing data and plot
2. Read DATA
3. Activities per day
4. Histogram of total number of steps/day
5. Mean and median
6. The 5-minute interval
7. Code
8. Histogram of total number of steps/day after imputting missing values
9. Panel plot comparing average number of steps/5-minute interval between weekends and weekdays

## 1. Parsing data and plot
```{r, echo=TRUE}
library(ggplot2)
library(lubridate)
library(dplyr)
```

## 2. Read DATA
```{r, echo=TRUE}
setwd("~/Dropbox/DATA SCIENCE/COURSE 5 WEEK 2 PROJECT")
activities <- read.csv("activity.csv")
activities$date <- ymd(activities$date)
```

## 3. Activities per day
```{r, echo=TRUE}
activities_per_day <- group_by(activities, date)
activities_per_interval <- group_by(activities, interval)
```

## 4. Histogram of total number of steps/day
```{r, echo=TRUE}
total <- tapply(activities$steps, activities$date, FUN=sum)
barplot(total, las=2, cex.names = 0.5)
```

## 5. Mean and median
```{r, echo=TRUE}
mean_median <- summarise(activities_per_day, mean=mean(steps, na.rm = TRUE), median=median(steps, na.rm = TRUE))
plot(mean_median$mean, type="l", col="red", ylab = "Quantity of Steps", xlab = "Number of Day", las=2, main="Average")
lines(mean_median$median, type="l", col="blue")
legend("topright", legend=c("mean", "median"), pch=1, col=c("red", "blue"))
```

## 6. The 5-minute interval

```{r, echo=TRUE}
average_of_steps <- summarise(activities_per_interval, average=mean(steps, na.rm = TRUE))
ranking_of_intervals <- arrange(average_of_steps, desc(average))
print(ranking_of_intervals$interval[1])
```

## 7. Code

```{r, echo=TRUE}
nas <- is.na(activities$steps)
total_empty <- sum(nas)
percentage_empty <- mean(nas)
# Total empty results
print(total_empty)

# Percentage
print(paste(sprintf("%.2f", percentage_empty * 100), "%", sep=""))
```

```{r, echo=TRUE}
not_empty <- mutate(activities)
na_intervals <- not_empty$interval[nas]
average_na_intervals <- sapply(na_intervals, function(item) {average_of_steps$average[average_of_steps$interval==item]})
not_empty$steps[nas] = average_na_intervals
#average_of_steps_by_interval$average[average_of_steps_by_interval$interval==0]
```


## 8. Histogram of total number of steps/day after imputting missing values

```{r, echo=TRUE}
not_empty_items <- tapply(not_empty$steps, not_empty$date, FUN=sum)

par(mfrow=c(2,1))
barplot(total, las=2, cex.names = 0.5, main="Histogram with NAs")
barplot(not_empty_items, las=2, cex.names = 0.5, main="Histogram without NAs")
```

# Missing values change our results.

## 9. Panel plot comparing average number of steps/5-minute interval between weekends and weekdays

```{r, echo=TRUE}
activities_with_days <- mutate(activities, weekday=weekdays(date))
activities_by_day_interval <- group_by(activities_with_days, weekday, interval)

results_per_day <- summarise(activities_by_day_interval, steps=mean(steps, na.rm=TRUE))
qplot(interval, steps, data=results_per_day, facets= weekday ~ .) + geom_line()
```
